 Reproducible Research: Peer Assessment 1


---
## Set up environment
```{r, echo=TRUE}
library(tidyr)
library(dplyr)
library(ggplot2)


```

## Loading and preprocessing the data
```{r, echo=TRUE}
unzip("./activity.zip")
steps <- read.csv("./activity.csv",header = TRUE, sep = ',', 
                  colClasses = c("numeric", "character", "integer"))
```

```{r, echo=TRUE}
library(lubridate)
steps$date <- as.Date(strptime(steps$date, format="%Y-%m-%d"))
str(steps)
head(steps)
```

## Whats is mean total number of steps take per day?
```{r, echo=TRUE}
stepsbyDay<- steps %>% filter(!is.na(steps)) %>%  group_by(date) %>%
  summarize(steps = sum(steps))
 head(stepsbyDay)
```

### Ploting means

``` {r}
ggplot(stepsbyDay, aes(x = steps)) +
  geom_histogram(col= "black", fill = "yellow", binwidth = 1000) +
  labs(title = "Histogram of Steps per day", x = "Steps per day", y = "Frequency")
```


### What is mean total number of steps taken per day?
```{r, echo=TRUE}
meanSteps <- steps %>% group_by(date) %>% summarise(meanstepsPerDay = mean(steps, na.rm= TRUE)) %>% summarise(meanStep= mean (meanstepsPerDay, na.rm= TRUE))

medianSteps <- steps %>% group_by(date) %>% summarise(medianStepsPerDay= median(steps, na.rm = TRUE)) %>% summarise(medianSteps= median(medianStepsPerDay, na.rm = TRUE))

medianSteps
meanSteps
```

## What is the average daily activity pattern?
```{r, echo=TRUE}
interval <- steps %>% group_by(interval) %>%  summarize(steps = mean(steps, na.rm = TRUE))
interval
```

### Ploting
```{r, echo=TRUE}

ggplot(interval, aes(x=interval, y=steps)) +
  geom_line(color = "firebrick")+ labs(title = "Averange daily activity pattern by 5 minutes intervals")+ xlab("Interval")+ ylab("Steps")
``` 

### Inerval with maximun of steps
```{r}
interval[which.max(steps$steps),]
```


## Imputing missing values 
```{r}
sum(is.na(interval))
```

### Filling NA with averange of 5 minutes interval
```{r}
fullSteps <- steps
Nas <- is.na(fullSteps$steps)
avg_interval <- tapply(fullSteps$steps, fullSteps$interval, mean, na.rm= TRUE, simplify = TRUE )
fullSteps$steps[Nas] <- avg_interval[as.character(fullSteps$interval[Nas])]
sum(is.na(fullSteps$steps))
```

### Recalculate the mean
```{r}
fullStepsNa <- fullSteps %>%
  filter(!is.na(steps)) %>%
  group_by(date) %>%
  summarize(steps = sum(steps)) %>%
  print
```

### Plot
```{r}
ggplot(fullStepsNa, aes(x = steps)) +
  geom_histogram(col= "black", fill = "yellow", binwidth = 1000) +
  labs(title = "Histogram of Steps per day. includin recalculates NA values", x = "Steps per day", y = "Frequency")
```

## Are there differences in activity patterns between weekdays and weekends?
```{r}
data_full <- mutate(fullSteps, weektype = ifelse(weekdays(fullSteps$date) == "Saturday" | weekdays(fullSteps$date) == "Sunday", "weekend", "weekday"))
data_full$weektype <- as.factor(data_full$weektype)
head(data_full)
```

### Ploting 
```{R}
interval_full <- data_full %>%
  group_by(interval, weektype) %>%
  summarise(steps = mean(steps))
s <- ggplot(interval_full, aes(x=interval, y=steps, color = weektype)) +
  geom_line() +
  facet_wrap(~weektype, ncol = 1, nrow=2)
print(s)
```